\papertitle

### 2020-01-28 superfri-template/paper.tex
Cleaning

\author{
%  Julian M. Kunkel
%  \textit{University of Reading}
%    \and
%  Luciana R. Pedro
%  \textit{University of Reading}
}
\date{\today}

@@

\section*{Abstract}
The efficient, convenient, and robust execution of data-driven workflows and enhanced data management are key for productivity in scientific computing.
Traditionally, in HPC, the concerns of storage and computing are separated and optimised independently from each other and the needs of the end-to-end user.
As climate and weather workflows become increasingly complex and blend beyond data centres while, at the same time, storage hierarchies become deeper, the community investigates means to re-organise storage access to utilise such environments fully.

@@

To satisfy the needs for different workflows, data centres are providing a infrastructure consisting of computing and storage devices with different characteristics which makes them more efficient for specific tasks.
Take, for example, the supercomputer Mistral at DKRZ that consists of 3,321 nodes.
It offers two types of compute nodes equipped with different CPUs and a range of GPU nodes.
Each node provides an SSD for local storage, and DKRZ has additionally two shared Lustre file systems with different performance characteristics.
Individual users and projects are mapped to one file system explicitly, and users can access it as work or scratch file system: While data is kept on the work file system indefinitely, its available space is limited by a quota.
The scratch file system allows to store more data, but data is automatically deleted after some time.

@@

\begin{minipage}{0.2\textwidth}
\begin{figure}[H]
  \includegraphics[width=\textwidth]{layers-xios}
  \caption{I/O Path for an MPI-Parallel Application}
  \label{fig:layers}
\end{figure}
\end{minipage}
\qquad
\begin{minipage}{0.7\textwidth}
In our example, we assume the application is parallelised for performance reasons and uses MPI to run the application.
It may use XIOS to gather data from individual fields which then uses NetCDF4 to store the data as a file.
XIOS provides a domain-level semantics and ...\jk{TODO}
Under the hood, NetCDF4 used the HDF5 API and file format.
Internally, HDF5 uses MPI again and its data types to specify the nature of the data stored.
Finally, data is stored on a parallel file system like Lustre, which, on the server-side, stores data in the local file system as block devices on storage media such as SSDs and HDDs.

@@

The appropriate software stack involved in executing a workflow is depicted in \Cref{fig:stages}.
As a preliminary, the user had specified a Cylc suite representing the workflow and provided scripts files for the individual tasks to execute; now the user enacts Cylc to start the workflow.
In the following, each stage of the execution is described further.

\begin{figure}[H]
  \centering
  \includegraphics[scale=1.4]{stages}
  \caption{Software Stack and Stages of Execution}
  \label{fig:stages}
\end{figure}

\begin{enumerate}
  \item \textbf{Cylc} Cylc parses the suite file, generate dependencies and defines a schedule for the execution.
  Cylc monitors the progress of the workflow.
  Once a task can be executed (dependencies are fulfilled), Cylc generates a Slurm script on the fly with the required metadata for Slurm that will invoke the user-defined script for that task and runs Slurm to queue the job.

  \item \textbf{Slurm} Schedules the execution of the job considering the scheduling policy of the data centre.
  Once the job is ready to be dispatched, i.e., resources are available, and the job priority is the highest, it is started on the supercomputer.

  \item \textbf{Job} The user-provided script is executed on a single node assigned to the job.
  The script has set environment variables containing information about the Cylc execution (e.g., a task in the workflow) and the Slurm job (e.g., number of nodes).

  \item \textbf{Script} The shell-script executes the sequence of commands, including applications that are parallel programs.
  Typically, to create a filename, the script calls Cylc commands with a template provided by the user and feeds this information to the application which, ultimately, defines the storage location to be used.
  A script typically consists of multiple fine-grained commands and may even call multiple parallel applications.

  \item \textbf{Application} The user-specified parallel application runs taking the filenames as defined in a configuration file or as arguments.
  The application may use NetCDF or XIOS to perform the I/O.
\end{enumerate}

@@

Consider the Cylc suite configuration for a toy monthly cycling workflow in \Cref{fig:cylc}.
In this workflow, a warm-cycled atmospheric model (\textbf{model}) simulates the physics from a current state to predict the future after a timestep, for example, after six hours.
This process is repeated in the model to simulate month and years into the future.
Once the simulation of any month is computed, the data for this month becomes available and can be analysed.
A user has specified a detailed workflow in a Cylc suite file (\texttt{suite.rc}).
In this workflow, the \textbf{model} is followed by postprocessing (\textbf{post}), forecast verification (\textbf{ver}), and product generation (\textbf{prod}) tasks.
Also, consider, in addition, a task \textbf{check} that compares some verification metric against products from two cycles earlier.

@@

What happens in many current workflows is that they use the only \textbf{work} and \textbf{scratch}.
When a task is set to computing, the corresponding dataset is moved to \textbf{scratch}, processed, and the resulting datasets are transferred back to \textbf{work}.
If the \textbf{scratch} filesystem achieves its capacity, then the datasets move back to \textbf{work} and keep running the task until it is finished.
If the I/O planning is made a priori, this situation can be avoided.
In addition to the file systems described, many supercomputers also have available for the user \textbf{scratch2} file systems (sometimes local and global).
Therefore, in a real scenario, if the architecture of the machine is known, we can better utilise its resources.

\Cref{fig:lifecycle} shows three possible life cycles and the access pattern for a specific dataset.
As shown in \Cref{fig:lifecycle}(a), the dataset could be stored on the \textbf{local} storage first to avoid congestion on the \textbf{work} file system, then be migrated to \textbf{work} file system where subsequent tasks of the workflow may read it multiple times.
In the end, this dataset might be an intermediate product that can then be deleted.
Alternatively (\Cref{fig:lifecycle}(b)), it could be stored on the \textbf{scratch} file system immediately and accessed there.
However, that would require that the last read access happens before files on \textbf{scratch} are automatically removed.
The last scenario (\Cref{fig:lifecycle}(c)) presents the case where data is created on \textbf{work} by a task. Still, then it is copied to a \textbf{local} node to allow read accesses of subsequent tasks which might be beneficial for small random accesses.
However, it would require that the subsequent tasks are placed on the same machine where data is now stored.
The scientist responsible for the workflow must optimise the available resources intuitively.
These are just examples, and currently, the scientists perform such kind of optimisations manually.

\begin{figure}[H]
  \centering
  \begin{subfigure}{.45\textwidth}
  \includegraphics[width=0.9\columnwidth]{lifecycle-1}
  \caption{Local and work}
  \end{subfigure}
  \begin{subfigure}{.45\textwidth}
  \includegraphics[width=0.9\columnwidth]{lifecycle-2}
  \caption{Scratch only}
  \end{subfigure}

@@

\cite{Vladimirov2014FileIO}

@@

Applied scientists should not spend time understanding hardware characteristics, but using their time to develop their work, let the computer do the job the most straightforward way as possible, and just collect and analyse the results.
However, nowadays, to be able to run a job in a HPC environment efficiently, researchers have to have profound knowledge about their workflow, which is expected, but also about decisions regarding storage, communication, and computing.

Data system do not respect the importance of data depends at all. Aspects like costs to reproduce the data (can it be reproduced easily), the type of the experiment (test run, production run) and runtime constraints for the overall and potential crucial workflow are not considered.
Value and priority should influence fault-tolerance strategies and imply the quality of service for performance and availability.

@@

Our vision requires two additional information: firstly, the user must augment the workflow description with information about I/O requirements and explicitly annotate dependencies to data sets; secondly, the data centre (or expert user) must provide a description of the storage architecture and characteristics for each storage system.

There are many approaches to implement the technology for this vision. It is apparent that changes need to be made to several software components to realise this vision.
In the next section, we will discuss the specific design for our transitional roadmap for climate and weather.

@@

Indeed, by monitoring I/O accesses, we can automatically draft such I/O descriptions on behalf of the user to simplify the specification.
All we need is to run one cycle of the specific workflow, considering that a workflow cycle is executed many times in climate and weather; this is a cheap approach.
From the recorded I/O accesses, we analyse the I/O patterns and dependencies associated with it to generate the configuration file.

@@

The system information shall comprise of all available storage systems, the system topology, and the characteristics of each of its required components.
This may include simplified or complex models of the components to approximate performance for specific I/O patterns allowing to predict performance behaviour.
It is expected that the data centre (or expert users) can create such a configuration file, for instance, by using vendor-provided information or by executing benchmarks.

\subsection{Smarter I/O Scheduling}

With this separation of concerns, we can abstract from the workflow what is essential and what a machine should optimise to ensure that the resources available are being used smartly.
In this paper, we focus on utilising a heterogeneous storage environment instead of exhaustively discussing the potential that is unleashed with the right level of abstraction in place by researchers that exploit the available opportunities.
Every time parameters are manually tuned right now there is a room from improvement by using automatised decision making.
And Machine Learning has been proven effective in replacing human-beings and adapt the workflow dynamically in case of unexpected situations (the system is down, nodes are unresponsive, etc.).

Optimisations we discuss in this paper are related to the life cycle of datasets and the placement of such datasets into specific storage according to system performance characteristics and the workflow specification.
Firstly, the I/O scheduler could make the placement decision for the individual data sets and manage their relocation during the life cycle, as discussed in \Cref{sec:datamanagement}.

Consider the hypothetical case to have just two tasks A and B that need to be repeated many times:
Datasets generated by Task\,\textbf{A} are used by Task\,\textbf{B} and hence both tasks must have access to the same data in the system.
Because there is one script responsible for setting the configuration, the decision in which directory to store data is somewhat fixed.
Such configuration was done manually and with restricted information about the storage and system architecture.
It would be interesting to explore storage options for the datasets and, e.g., have different cycles stored placed in different storage systems.
If the lifetime of such dataset is, for instance, two cycles, alternate the datasets into two scratch file systems is something that would be a simple job an I/O-aware scheduler.
However, manually, that implies at having at least two scripts\footnote{Complicated scripts would have allowed changing the storage type depending on the cycle. Still, it is a significant burden to the user.} for each task with information about the different storage.

\Cref{fig:cycle4} exemplifies a restriction of our example climate/weather workflow the way it was being generated.
Assume you have, for instance, four cycles of the model running with all the necessary post-processing steps.
Typically, the user has written the scripts for each task by defining a filename using the same prefix that is then placed at the same storage type every time.
That happens because all datasets generated by Task \textbf{model} will be used by Task \textbf{post\,prod}, so they should stay on the same node.
The user-specific storage locations are a viable option to run a workflow.
The user might even have analysed various scenarios such as depicted in
\Cref{sec:datamanagement}.

Additionally, by knowing how to rerun the workflow behind the data creation, a smarter storage system could trade data security with potential recomputation opportunities to optimise the cost-efficiency of the system.
Intermediate states could be rerun by utilising virtualisation and container technologies.
Similarly, the scheduler may decide to couple subsequent steps of a workflow directly without storing intermediate data products on persistent storage.
Finally, data might be replicated but enables the system to rerun parts of the workflow in case of a data loss.

@@

Once we have the complete knowledge about how the workflow is expected to run, we can optimise steps that were before predefined by the user, usually considering only options for feasibility.

@@

\item Optimisation: The workflow will be optimised for the specific available system characteristics considering potentially runtime characteristics and storage resilience which will increase performance and system utilisation.
By knowing the workflow behind the data creation, the system could trade data security with potential recomputation opportunities to optimise the cost-efficiency.

\item Performance-portability: The user has to specify the I/O requirements only once for the tasks of the workflow.
The workflow can now be run on any system without user intervention.
Even more, if the system characteristics change, e.g., it gets upgraded, and an additional storage tier becomes available, or if storage degrades, the I/O-aware workflow could automatically adapt and make use of this new environment.

@@

This section describes our approach to incrementally extend the current software stack for climate and weather workflows that realises parts our vision.

To automatically make scheduling decisions, the software stack needs to:
\begin{enumerate}
\item delivers information about information life cycle together with the workflow, and
\item adapts the resulting workflow, individual script and application execution to consider the automatic I/O schedule.
\end{enumerate}

@@

\Cref{fig:cycle1} introduces one cycle of the suite in \Cref{fig:cylc} with a simplified notation. The arrows represent the tasks dependencies, but there is no information about the data that are generated. Tasks \textbf{verification} and \textbf{post\,proc} depend on Task \textbf{model}, but does Task \textbf{model} needs to be fully completed to satisfy the dependency?
Of course, once Task \textbf{model} is completed, the dependent tasks can be executed, and that is the type of information Cylc has to create and reproduce the workflow.

\begin{figure}[H]
  \begin{subfigure}{0.45\textwidth}
    \centering
    \includegraphics[width=\columnwidth]{cycle1}
    \caption{Original}
    \label{fig:cycle1}
  \end{subfigure}
  \qquad
  \begin{subfigure}{0.45\textwidth}
    \centering
    \includegraphics[width=\columnwidth]{cycle-io-dep}
    \caption{Including datasets and dependencies}
    \label{fig:cycle-io-dep}
  \end{subfigure}

@@

The scenario presented in \Cref{fig:cycle-io-dep} shows the same cycle of \Cref{fig:cycle1}, but  extended information about datasets is added.
In this picture, we have the tasks dependencies, and we explicitly list the datasets dependencies regarding the tasks that generated them.
The way data moves through tasks is by creating datasets to pass on the information, including filename and storage destination.
The datasets produced by every task are coded in the workflow scripts before Cylc evaluation, and it is just replicated by this tool. The tasks can create a different number of datasets, and these datasets may be necessary for the immediate next task in the current cycle, or tasks in different cycles.

@@

The user can make decisions about storage, and the processing of the datasets/tasks and s/he will insert this information as an additional file to the Cylc workflow management.
In fact, the user is already doing that, but he is doing that in just one specific way, to make the workflow possible. Here, we are thinking about simplifying the information the user is already providing. So, instead of giving the complete workflow to Cylc, we are expecting the user to provide straightforward information, like:

@@

Now, comes the question. Assuming the information is available, who will be responsible for making the decisions for optimising the storage and the processing tasks? The answer here is simple: the ESDM Scheduler will do it. The ESDM Scheduler can be used in different steps of the process. ESDM has already an interface with NetCDF. Assuming the user can provide extra information in advance, ESDM can interact in the process before the user gives the data to the Cylc tool, using the information to construct an optimised workflow. Once Cylc builds the workflow, ESDM can interact again after the configuration of the scripts for both effective and efficient run of the parallel applications regarding the available storage and architecture.

@@

The appropriate software stack involved in executing a workflow with our I/O-aware scheduler is depicted in \Cref{fig:stages-io}.
In the following, we describe the modifications to the current execution (described in \Cref{fig:stages}) we propose in this vision paper.
Firstly, the end-user now has to provide an additional file that covers the I/O information for each task; slight changes have to be made to the scripts that are executed; these will be discussed below.

\begin{figure}[H]
  \centering
  \includegraphics[scale=1.4]{stages-io}
  \caption{Software Stack and Stages of Execution with our I/O-Aware Scheduler}
  \label{fig:stages-io}
\end{figure}

\begin{enumerate}
  \item \textbf{Cylc}\ \ Cylc parses the suite file, generate dependencies and defines a schedule for the execution.
  Now, Cylc calls EIOS to identify potential to optimise the schedule.
  EIOS may have decided that subsequent jobs shall be placed on the same node or to reorder the execution of some jobs.
  Also, it may have provided information for conducting data migration.

  \item \textbf{EIOS}\ \
  Reads information the workflow: the original suite file and the I/O suit file.
  Acts depending on the stage of execution, i.e., we could be preparing execution in Cylc, or we are already inside a script.
  It loads the \texttt{esdm.conf} file that contains the information about the available storage technology in the data centre and their characteristics.
  Provides the information on how to prioritise I/O targets.

@@

  \item \textbf{Slurm}\ \ Schedules the execution of the job considering the scheduling policy of the data centre.
  Cylc may now have added a constraint to place a job on the same nodes as a previous job which is handled by the modified Slurm.
  Also, if migrations have to be performed before or after a job is run, Slurm will administer them.

@@

  The script has set environment variables containing information about the Cylc execution (e.g., a task in the workflow) and the Slurm job (e.g., number of nodes).

@@

  Typically, to create a filename, the script calls Cylc commands with a template provided by the user and feeds this information to the application.
  This ultimately defines the storage location to be used.
  A script typically consists of multiple fine-grained commands and may even call multiple parallel applications.

@@

\section{Extra Figures}

\begin{figure}[H]
  \centering
  \includegraphics[scale=0.6]{workflow-all}
  \caption{Workflow All.}
  \label{fig:work}
\end{figure}

From the user perspective, a general workflow can be described by three categories described in \Cref{tab:workflow}.

Description & information inserted in the workflow. & Tasks specified by the user. & Possible outcomes of the tasks. \\ \hline

For this work, Cylc is an essential tool regarding not only the execution of the workflow but the exchange of information with the EIOS about I/O and storage.
Cylc receives, as predefined, information about the storage and dependencies of the datasets. The current information is used to construct and run the complete workflow, but it is not employed in optimising intermediate steps.
The configuration is already predefined, and it cannot change. For instance, if it is established that Task\,2 depends on Task\,1, then we have to wait for Task\,1 to be completed to start running Task\,2.
Thinking what about I/O, no decision is being made about an optimal way to store the data.
Consider, as an example, that the datasets from \texttt{product1.ds} (the final product of task 1) are divided into \texttt{product1-final.ds} and \texttt{product1-task2.ds}. Then, it is clear that \texttt{product1-task2.ds} will be required to run Task\,2, and, because of that, should be put in fast storage to quickly being retrieved, or even kept in local storage. The \texttt{product1-final.ds} dataset, however, can go directly to tape, if the lifetime of the final output has to be stored for a long time, but no access is required from the current workflow.
The bottom line here is: several options can be optimised if we know the whole workflow in advance.

@@

>>>>

\author{Julian M. Kunkel
  \textit{University of Reading}
    \and
  Luciana R. Pedro
  \textit{University of Reading}
}
\date{\today}

@@

\section*{Abstract}
The efficient, convenient, and robust execution of data-driven workflows and enhanced data management are key for productivity in scientific computing.
Traditionally, in HPC, the concerns of storage and computing are separated and optimised independently from each other and from the needs of end-to-end user.
As climate and weather workflows become increasingly complex and blend beyond data centres while, at the same time, storage hierarchies become deeper, the community investigates means to re-organise storage access to utilise such environments fully.

@@

To satisfy the needs for different workflows, data centres are providing a infrastructure consisting of computing and storage devices with different characteristics which makes them more efficient for specific tasks.
Take, for example, the supercomputer Mistral at DKRZ that consists of 3,321 nodes.
It offers two types of compute nodes equipped with different CPUs and a range of GPU nodes.
Each node provides an SSD for local storage and DKRZ has additionally two shared Lustre file systems with different performance characteristics.
Individual users and projects are mapped to one file system explicitly and users can access it as work or scratch file system: While data is kept on the work file system indefinitely, its available space is limited by a quota.
The scratch file system allows to store more data, but data is automatically deleted after some time.

@@

\begin{minipage}{0.2\textwidth}
\begin{figure}[H]
  \includegraphics[width=\textwidth]{layers-xios}
  \caption{I/O Path for an MPI-Parallel Application}
  \label{fig:layers}
\end{figure}
\end{minipage}
\qquad
\begin{minipage}{0.7\textwidth}
In our example, we assume the application is parallelised for performance reasons and uses MPI to run the application.
It may use XIOS to gather data from individual fields which then uses NetCDF4 to store the data as a file.
XIOS provides a domain-level semantics and ...\jk{TODO}
Under the hood, NetCDF4 used the HDF5 API and file format.
Internally, HDF5 uses MPI again and its data types to specify the nature of the data stored.
Finally, data is stored on a parallel file system like Lustre, which, on the server-side, stores data in local file system as block devices on storage media such as SSDs and HDDs.

@@

The relevant software stack involved in executing a workflow is depicted in \ref{fig:stages}.
As a preliminary, the user had specified a Cylc suite representing the workflow and provided scripts files for the individual tasks to execute; now the user enacts Cylc to start the workflow.
In the following, each stage of the execution is described further.

\begin{figure}[bt]
  \centering
  \includegraphics[scale=1.4]{stages}
  \caption{Software Stack and Stages of Execution}
  \label{fig:stages}
\end{figure}

\begin{enumerate}
  \item \textbf{Cylc}\ \ Cylc parses the suite file, generate dependencies and defines a schedule for the execution.
  Cylc monitors the progress of the workflow.
  Once a task can be executed (dependencies are fulfilled), Cylc generates a Slurm script on the fly with the required metadata for Slurm that will invoke the user-defined script for that task, and runs Slurm to queue the job.

  \item \textbf{Slurm}\ \ Schedules the execution of the job considering the scheduling policy of the data center.
  Once the job is ready to be dispatched, i.e., resources are available and the job priority is the highest, it is started on the supercomputer.

  \item \textbf{Job}\ \ The user-provided script is executed on a single node assigned to the job.
  The script has set environment variables containing information about the Cylc execution (e.g., task in the workflow) and Slurm job (e.g., number of nodes).

  \item \textbf{Script}\ \ The shell-script executes the sequence of commands including applications that are parallel programs.
  Typically, to create a filename, the script calls Cylc commands with a template provided by the user and feeds this information to the application.
  This ultimately defines the storage location to be used.
  A script typically consist of multiple fine-grained commands and may even call multiple parallel applications.

  \item \textbf{Application}\ \ The user-specified parallel application runs taking the filenames as defined in a configuration file or as arguments.
  The application may use NetCDF or XIOS to perform the I/O.
\end{enumerate}

@@

Consider the Cylc suite configuration for a toy monthly cycling workflow in Figure \ref{fig:cylc}.
In this workflow, a warm-cycled atmospheric model (\textbf{model}) simulates the physics from a current state to predict the future after a timestep, for example, after six hours.
This process is repeated in the model to simulate month and years into the future.
Once the simulation of any month is computed, the data for this month becomes available and can be analysed.
A user has specified a detailed workflow in a Cylc suite file (\texttt{suite.rc}).
In this workflow, the \textbf{model} is followed by postprocessing (\textbf{post}), forecast verification (\textbf{ver}), and product generation (\textbf{prod}) tasks.
Also, consider, in addition, a task \textbf{check} that compares some verification metric against products from two cycles earlier.

@@

What happens in many current workflows is that they use only \textbf{work} and \textbf{scratch}.
When a task is set to computing, the corresponding dataset is moved to \textbf{scratch}, processed, and the resulting datasets are transferred back to \textbf{work}.
If the \textbf{scratch} filesystem achieves its capacity, then the datasets move back to \textbf{work} and keep running the task until it is finished.
If the I/O planning is made a priori, this situation can be avoided.
In addition to the file systems described, many supercomputers also have available for the user \textbf{scratch2} file systems (sometimes local and global).
Therefore, in a real scenario, if the architecture of the machine is known, we can better utilize its resources.

\Cref{fig:lifecycle} shows three possible life cycles and the access pattern for a specific dataset.
As shown in \Cref{fig:lifecycle}(a), the dataset could be stored on the \textbf{local} storage first to avoid congestion on the \textbf{work} file system, then be migrated to \textbf{work} file system where subsequent tasks of the workflow may read it multiple times.
In the end, this dataset might be an intermediate product that can then be deleted.
Alternatively (\Cref{fig:lifecycle}(b)), it could be stored on the \textbf{scratch} file system immediately and accessed there.
However, that would require that the last read access happens before files on \textbf{scratch} are automatically removed.
The last scenario (\Cref{fig:lifecycle}(c)) presents the case where data is created on \textbf{work} by a task, but then it is copied to a \textbf{local} node to allow read accesses of subsequent tasks which might be beneficial for small random accesses.
However, it would require that the subsequent tasks are placed on the same machine where data is now stored.
The scientist responsible for the workflow must optimise the available resources intuitively.
These are just examples and currently the scientists perform such kind of optimisations manually.

\begin{figure}[tb]
  \centering
  \begin{subfigure}{.45\textwidth}
  \includegraphics[width=0.9\columnwidth]{lifecycle-1}
  \caption{Local and work}
  \end{subfigure}
  \begin{subfigure}{.45\textwidth}
  \includegraphics[width=0.9\columnwidth]{lifecycle-2}
  \caption{Scratch only}
  \end{subfigure}

@@

@@

Applied scientists should not spend time understanding hardware characteristics, but using their time to develop their work, let the computer do the job the simplest way as possible, and just collect and analyse the results.
However, nowadays, to be able to run a job in a HPC environment efficiently, researchers have to have profound knowledge about their workflow, which is expected, but also about decisions regarding storage, communication, and computing.

Data system do not respect the importance of data depends at all; aspects like costs to reproduce the data (can it be reproduced easily), the type of the experiment (test run, production run) and runtime constraints for the overall and potential crucial workflow are not considered.
Value and priority should influence fault-tolerance strategies and imply the quality of service for performance and availability.

@@

Our vision requires two additional information: firstly, the user must augment the workflow description with information about IO requirements and explicitly annotate dependencies to data sets; secondly, the data center (or expert user) must provide a description of the storage architecture and characteristics for each storage system.

There are many approaches to implement the technology for this vision, it is apparent that changes need to be made to several software components to realize this vision.
In the next section, we will discuss the specific design for our transitional roadmap for climate and weather.

@@

Indeed, by monitoring I/O accesses we can automatically draft such IO descriptions on behalf of the user to simplify the specification.
All we need is to run one cycle of the specific workflow, considering that a workflow cycle is executed many times in climate and weather, this is a cheap approach.
From the recorded IO accesses, we analyze the IO patterns and dependencies associated with it to generate the configuration file.

@@

The system information shall comprise of all available storage systems, the system topology, and the characteristics of each of the involved components.
This may include simplified or complex models for the component to approximate performance for specific I/O patterns allowing to predict performance behavior.
It is expected that the data center (or expert users) can create such a configuration file, for instance, by using vendor provided information or by executing benchmarks.

\subsection{Smarter IO Scheduling}

With this separation of concerns, we can abstract from the workflow what is essential and what a machine should optimise to ensure that the resources available are being used smartly.
In this paper, we focus on utilizing a heterogeneous storage environment instead of exhaustively discussing the potential that is unleashed with the right level of abstraction in place by researchers that exploit the available opportunities.
Every time parameters are manually tuned right now is room from improvement by using automatised decision making.
And Machine Learning has been proven effective in replacing human-beings and adapt the workflow dynamically in case of unexpected situations (the system is down, nodes are unresponsive, etc.).

Optimisations we discuss in this paper are related to the life cycle of datasets and the placement of such datasets into specific storage according to system performance characteristics and the workflow specification.
Firstly, the IO scheduler could make the placement decision for the individual data sets and manage their relocation during the life cycle as discussed in \Cref{sec:datamanagement}.

Consider the theoretical case to have just two tasks A and B that need to be repeated many times:
Datasets generated by Task\,\textbf{A} are used by Task\,\textbf{B} and hence both tasks must have access to the same data in the system.
Because there is one script responsible for setting the configuration, the decision in which directory to store data is somewhat fixed.
Such configuration was done manually and with restricted information about the storage and system architecture.
It would be interesting to explore storage options for the datasets and, e.g., have different cycles stored placed in different storage systems.
If the lifetime of such dataset is, for instance, two cycles, alternate the datasets into two scratch file systems is something that would be a simple job an I/O-aware scheduler.
However, manually, that implies at having at least two scripts\footnote{Complicated scripts would have allowed to change the storage type depending on the cycle but it is big burden to the user.} for each task with information about the different storage.

Figure \ref{fig:cycle4} exemplifies a restriction of our example climate/weather workflow the way it was being generated.
Assume you have, for instance, four cycles of the model running with all the required post-processing steps.
Typically, the user has generated the scripts for each task by generating a filename using the same prefix that is then placed at the same storage type every time.
That happens because all datasets generated by Task \textbf{model} will be used by Task \textbf{post\,prod}, so they should stay on the same node.
The user-specific storage locations are a viable option to run a workflow.
The user might even have analysed various scenarios such as depicted in
\Cref{sec:datamanagement}.

Additionally, by knowing how to rerun the workflow behind the data creation, a smarter storage system could trade data security with potential recomputation opportunities to optimize the cost-efficiency of the system.
Intermediate states could be rerun by utilising virtualisation and container technologies.
Similarly, the scheduler may decide to couple subsequent steps of a workflow directly without storing intermediate data products on persistent storage.
Finally, data might be replicated but enables the system to rerun parts of the workflow in case of a data loss.

@@

Once we have the complete knowledge about how the workflow is expected to run, we can optimise steps that were before predefined by the user usually considering only options for feasibility.

@@

\item Optimization: The workflow will be optimised for the specific available system characteristics considering potentially even runtime characteristics and storage resilience.
This will increase performance and system utilization.
By knowing the workflow behind the data creation, the system could trade data security with potential recomputation opportunities to optimize the cost-efficiency.

\item Performance-portability: The user has to specify the IO requirements only once for the tasks of the workflow.
The workflow can now be run on any system without user intervention.
Even more, if the system characteristics change, e.g.. it gets upgraded and an additional storage tier becomes available or if storage degrades, the I/O-aware workflow could automatically adapt and make use of this new environment.

@@

This section describes our approach to incrementally extent the current software stack for climate and weather workflows that realizes parts our vision.

In order to automatically make scheduling decisions the software stack needs to:
\begin{enumerate}
\item deliver information about information life cycle together with the workflow, and
\item adapt the resulting workflow, individual script and application execution to consider the automatic IO schedule.
\end{enumerate}

@@

Figure \ref{fig:cycle1} introduces one cycle of the suite in Figure \ref{fig:cylc} with a simplified notation. The arrows represent the tasks dependencies, but there is no information about the data that are generated. Tasks \textbf{verification} and \textbf{post\,proc} depend on Task \textbf{model}, but does Task \textbf{model} needs to be fully completed to satisfy the dependency?
Of course, once Task \textbf{model} is completed, the dependent tasks can be executed, and that is the type of information Cylc has to create and reproduce the workflow.

\begin{figure}[bt]
  \begin{subfigure}{0.45\textwidth}
    \centering
    \includegraphics[width=\columnwidth]{cycle1}
    \caption{Original}
    \label{fig:cycle1}
  \end{subfigure}
  \qquad
  \begin{subfigure}{0.45\textwidth}
    \centering
    \includegraphics[width=\columnwidth]{cycle-io-dep}
    \caption{Including datasets and dependencies}
    \label{fig:cycle-io-dep}
  \end{subfigure}

@@

The scenario presented in \Cref{fig:cycle-io-dep} shows the same cycle of Figure \ref{fig:cycle1}, but  extended information about datasets is added.
In this picture, we have the tasks dependencies, and we explicitly list the datasets dependencies regarding the tasks that generated them.
The way data moves through tasks is by creating datasets to pass on the information, including filename and storage destination.
The datasets produced by every task are coded in the workflow scripts before Cylc evaluation, and it is just replicated by this tool. The tasks can create a different number of datasets, and these datasets may be necessary for the immediate next task in the current cycle, or tasks in different cycles.

@@

The user to be able to make decisions about storage and the processing of the datasets/tasks.
user will insert this information as an additional file to the Cylc workflow management.
In fact, the user is already doing that, but he is doing that in just one specific way, to make the workflow possible. Here, we are thinking about simplifying the information the user is already providing. So, instead of giving the complete workflow to Cylc, we are expecting the user to provide straightforward information, like:

@@

Now, comes the question. Assuming the information is available, who will be responsible for making the decisions for optimising the storage and the processing tasks? The answer here is simple: the ESDM Scheduler will do it. The ESDM Scheduler can be used in different steps of the process. ESDM has already an interface with NetCDF. Assuming the user can provide extra information in advance, ESDM can interact in the process before the user give the data to the Cylc tool, using the information to construct an optimised workflow. Once Cylc builds the workflow, ESDM can interact again after the configuration of the scripts for both effective and efficient run of the parallel applications regarding the available storage and architecture.

@@

The relevant software stack involved in executing a workflow with our IO-aware scheduler is depicted in \Cref{fig:stages-io}.
In the following, we describe the modifications to the current execution (described in \Cref{fig:stages}) we propose in this vision paper.
Firstly, the end-user now has to provide an additional file that covers the IO information for each task; slight changes have to be made to the scripts that are executed, these will be discussed below.

\begin{figure}[tb]
  \centering
  \includegraphics[scale=1.4]{stages-io}
  \caption{Software Stack and Stages of Execution with our IO-Aware Scheduler}
  \label{fig:stages-io}
\end{figure}

\begin{enumerate}
  \item \textbf{Cylc}\ \ Cylc parses the suite file, generate dependencies and defines a schedule for the execution.
  Now, Cylc calls EIOS to identify potential to optimize the schedule.
  EIOS may have decided that subsequent jobs shall be placed on the same node or to reorder the execution of some jobs.
  Also it may have provided information for conducting data migration.

  \item \textbf{EIOS}\ \
  Reads information the workflow: the original suite file and the IO suit file.
  Acts depending on the stage of execution, i.e., we could be preparing execution in Cylc or we are already inside a script.
  It loads the \texttt{esdm.conf} file that contains the information about the available storage technology in the data center and their characteristics.
  Provides the information how to prioritize IO targets.

@@

  \item \textbf{Slurm}\ \ Schedules the execution of the job considering the scheduling policy of the data center.
  Cylc may now have added a constraint to place a job on the same nodes as a previous job which is handled by the modified Slurm.
  Also, if migrations have to be performed before or after a job is run, Slurm will administer them.

@@

  The script has set environment variables containing information about the Cylc execution (e.g., task in the workflow) and Slurm job (e.g., number of nodes).

@@

  Typically, to create a filename, the script calls Cylc commands with a template provided by the user and feeds this information to the application.
  This ultimately defines the storage location to be used.
  A script typically consist of multiple fine-grained commands and may even call multiple parallel applications.

@@

@@

\begin{figure}[H]
  \centering
  \includegraphics[scale=0.6]{workflow-all}
  \caption{Workflow All.}
  \label{fig:work2}
\end{figure}

From the user perspective, a general workflow can be described by three categories described in Table \ref{tab:workflow}.

Description & Information inserted in the workflow. & Tasks specified by the user. & Possible outcomes of the tasks. \\ \hline

For this work, Cylc is the essential tool regarding not only the execution of the workflow, but the exchange of information with the EWIOScheduler about I/O and storage.
Cylc receives, as predefined, information about the storage and dependencies of the datasets. The current information is used to construct and run the complete workflow, but it is not employed in optimising intermediate steps.
The configuration is already predefined, and it cannot change. For instance, if it is established that Task\,2 depends on Task\,1, then we have to wait for Task\,1 to be completed to start running Task\,2.
Thinking what about I/O, no decision is being made about an optimal way to store the data.
Consider, as an example, that the datasets from \texttt{product1.ds} (the final product of task 1) are divided into \texttt{product1-final.ds} and \texttt{product1-task2.ds}. Then, it is clear that \texttt{product1-task2.ds} will be required to run Task\,2, and, because of that, should be put in fast storage to quickly being retrieved, or even kept in local storage. The \texttt{product1-final.ds} dataset, however, can go directly to tape, if the lifetime of the final output has to be stored for a long time, but no access is required from the current workflow.
The bottom line here is: several options can be optimised if we know the whole workflow in advance.
